Temporal ordering and registration of images in studies of developmental dynamics

### 2015-02-11 image_analysis_paper_development2/manuscript.tex
added stas edits

\author{Carmeline~J.~Dsilva \thanks{Department of Chemical and Biological Engineering, Princeton University, Princeton, New Jersey, USA},
Bomyi~Lim \footnotemark[1],
Hang~Lu \thanks{School of Chemical and Biomolecular Engineering, Georgia Institute of Technology, Atlanta, Georgia, USA},
Amit~Singer \thanks{Department of Mathematics, Princeton University, Princeton, New Jersey, USA} \thanks{Program in Applied and Computational Mathematics, Princeton University, Princeton, New Jersey, USA}, Ioannis~G.~Kevrekidis \footnotemark[1] \footnotemark[4] ,\\ and
Stanislav~Y.~Shvartsman \footnotemark[1] \thanks{Lewis-Sigler Institute for Integrative Genomics, Princeton University, Princeton, New Jersey, USA}}

@@

We used vector diffusion maps to register and order the images.
%
\fig~\ref{fig:drosophila_live_imaging}B shows the images from \fig~\ref{fig:drosophila_live_imaging}A, now registered and ordered; the real time for each frame is also indicated.
%
With a small number of exceptions, the recovered ordering is consistent with the real time dynamics.
%
\fig~\ref{fig:drosophila_live_imaging}C and \fig~\ref{fig:drosophila_live_imaging}D  show the correlations between the recovered and true angles and rank orders, respectively, for the entire data set.
%
Both the angles and the ranks are recovered with a high degree of accuracy.

@@

%
Vector diffusion maps allow us to automatically register images, an essential task for many applications.
%
Simultaneously, the algorithm provides us with parameters to describe each image.
%
In the examples presented here, we have focused on ordering the images in time using the first vector diffusion maps coordinate.
%
In general, we can recover several coordinates which concisely and comprehensively describe the data set.
%
This parametrization can then be used for typical data analysis tasks, such as outlier detection and model fitting.
%
%Data which comes from several different genetic backgrounds can be clustered according to their spatial expression patterns, and clustering can also be used to separate images into different developmental stages.
%
Furthermore, images taken from different viewing directions can be analyzed, as the vector diffusion maps parametrization will organize the images according to the viewing angle \citep{singer2011viewing}.
%
Another direction for future work is related to the joint analysis of data sets provided by different imaging approaches, such as merging live imaging data of tissue morphogenesis with snapshots of cell signaling and gene expression from fixed embryos \citep{krzic2012multiview, ichikawa2014live, rubel2010coupling, dsilva2013nonlinear}.
%
In the future, it would also be interesting to explore the connections between our proposed approach and recently developed methods for ordering and classification of face images \citep{kemelmacher2011exploring, kemelmacher2014illumination}
%
Given the rapidly increasing volumes of imaging data from studies of multiple developmental systems, we expect that dimensionality reduction approaches discussed in this work will be increasingly useful for biologists and motivate future applications and algorithmic advances.

@@

Images were subsampled, normalized, blurred, and centered prior to diffusion maps analysis to remove any variations due to the experimental and imaging framework.
%
Details about the preprocessing operations applied to each imging data set are given in \SI.

>>>>

\author{Carmeline~J.~Dsilva \thanks{Department of Chemical and Biological Engineering, Princeton University, Princeton, New Jersey, USA},
Bomyi~Lim \footnotemark[1], \\
Hang~Lu \thanks{School of Chemical and Biomolecular Engineering, Georgia Institute of Technology, Atlanta, Georgia, USA},
Amit~Singer \thanks{Department of Mathematics, Princeton University, Princeton, New Jersey, USA} \thanks{Program in Applied and Computational Mathematics, Princeton University, Princeton, New Jersey, USA}, Ioannis~G.~Kevrekidis \footnotemark[1] \footnotemark[4] ,and
Stanislav~Y.~Shvartsman \footnotemark[1] \thanks{Lewis-Sigler Institute for Integrative Genomics, Princeton University, Princeton, New Jersey, USA}}

@@

We used vector diffusion maps to register and order the images.
%
\fig~\ref{fig:drosophila_live_imaging}B shows the images from \fig~\ref{fig:drosophila_live_imaging}A, now registered and ordered; the real time for each frame is also indicated.
%
With a small number of exceptions, the recovered ordering is consistent with the real time dynamics.
%
\fig~\ref{fig:drosophila_live_imaging}C and \fig~\ref{fig:drosophila_live_imaging}D  show the correlations between the recovered and true angles and rank orders, respectively, for the entire data set.
%
Both the angles and the ranks are recovered with a high degree of accuracy.
%
We note that determining which end of the trajectory corresponds to early in the developmental progression is a post-processing task that requires some {\em a priori} information.

@@

%
Vector diffusion maps allow us to automatically register images, an essential task for many applications.
%
Simultaneously, the algorithm provides us with parameters to describe each image.
%
In the examples presented here, we have focused on ordering the images in time using the first vector diffusion maps coordinate.
%
In general, we can recover several coordinates which concisely and comprehensively describe the data set.
%
This parametrization can then be used for typical data analysis tasks, such as outlier detection and model fitting.
%
%Data which comes from several different genetic backgrounds can be clustered according to their spatial expression patterns, and clustering can also be used to separate images into different developmental stages.
%
Furthermore, images taken from different viewing directions can be analyzed, as the vector diffusion maps parametrization will organize the images according to the viewing angle \citep{singer2011viewing}.
%
Another direction for future work is related to the joint analysis of data sets provided by different imaging approaches, such as merging live imaging data of tissue morphogenesis with snapshots of cell signaling and gene expression from fixed embryos \citep{krzic2012multiview, ichikawa2014live, rubel2010coupling, dsilva2013nonlinear}.
%
It would also be interesting to explore the connections between our proposed approach and recently developed methods for ordering and classification of face images \citep{kemelmacher2011exploring, kemelmacher2014illumination}
%
Given the rapidly increasing volumes of imaging data from studies of multiple developmental systems, we expect that dimensionality reduction approaches discussed in this work will be increasingly useful for biologists and motivate future applications and algorithmic advances.

@@

Images were subsampled, normalized, blurred, and centered prior to diffusion maps analysis to remove any variations due to the experimental and imaging framework.
%
Details about the specific preprocessing operations applied to each imging data set are given in \SI.
